---
title: "Exercise 3-1"
output: html_document
---
We begin by building the text messages data set based on the provided data. First the teen data:
```{r}
n = 800-696
teen_ones = rep(1,696)
teen_zeros = rep(0,n)
teen = c(teen_zeros, teen_ones)
```

Now the adult data:
```{r}
n = 2252-1621
adult_ones = rep(1,1621)
adult_zeros = rep(0,n)
adult = c(adult_zeros, adult_ones)
```

Here I define a function for bootstrapping a difference of proportions
```{r}
bootstrap_props = function(data1, data2, samples){
  prop_difs = numeric(samples)
  for (i in samples){
    data1_sample = sample(data1, size = length(data1), replace=TRUE)
    data2_sample = sample(data2, size = length(data2), replace=TRUE)
    prop1 = mean(data1_sample)
    prop2 = mean(data2_sample)
    prop_difs[i] = prop1-prop2
  }
  return(prop_difs)
}
```

Now we can perform the bootstrap and plot it
```{r}
prop_difs = bootstrap_props(teen, adult, 10000)
hist(prop_difs)
```

Now we can prepare the pregnant data set
```{r}
n = 135-38
p_smoker_ones = rep(1,38)
p_smoker_zeros = rep(0,n)
p_smoker = c(p_smoker_zeros, p_smoker_ones)
```
```{r}
n = 543-206
p_nonsmoker_ones = rep(1,206)
p_nonsmoker_zeros = rep(0,n)
p_nonsmoker = c(p_nonsmoker_zeros, p_nonsmoker_ones)
```

Performing the bootstrap
```{r}
prop_difs = bootstrap_props(p_smoker, p_nonsmoker, 10000)
hist(prop_difs)
```

Loading the Restaurant Tips data set and attaching it
```{r}
tips = read.csv("restaurantTipsBill.csv", header=TRUE)
attach(tips)
```

Here is a bootstrap for the correlation coefficients
```{r}
N=10^4
n=nrow(tips)
corrs = numeric(N)
for (i in 1:N){
  index = sample(n, replace=TRUE)
  tips_boot = tips[index, ]
  corrs[i] = cor(tips_boot["Tip"], tips_boot["Bill"])
}
hist(corrs)
```

Here is a bootstrap for the regression slopes
```{r}
N=10^4
n=nrow(tips)
slopes = numeric(N)
for (i in 1:N){
  index = sample(n, replace=TRUE)
  tips_boot = tips[index, ]
  regression = lm(tips_boot$Tip~tips_boot$Bill, data=tips_boot)
  slopes[i] = coef(regression)[2]
}
hist(slopes)
```




