---
title: '4'
output: html_document
---
First we import the data and examine the dataset
```{r}
library(MASS)
attach(Boston)
?Boston
```

We can provide a simple point estimate of the mean of medv which we call $\hat\mu$.
```{r}
medv_mean = mean(Boston[["medv"]])
medv_mean
```

We can estimate the standard error of $\hat\mu$, denoted $SE_{\hat\mu}$ via $SE_{\hat\mu}=S/\sqrt{n}$ where $S$ is the sample standard deviation.
```{r}
medv_sd = sd(Boston[["medv"]])
n = length(Boston[["medv"]])
mean_se = medv_sd/sqrt(n)
mean_se
```

Now we can use a bootstrap to estimate the standard error of $\hat\mu$
```{r}
N=10^4
mean_se_boot = numeric(N)
for (i in 1:N){
  medv_boot = sample(Boston[["medv"]], n, replace=TRUE)
  mean_se_boot[i] = sd(medv_boot)/sqrt(n)
}
mean(mean_se_boot)
```

This result is in very good agreement with the estimate above. We can create a 95% confidence interval using 
$[\hat\mu-2SE_{\hat\mu}, \hat\mu+2SE_{\hat\mu}]$
```{r}
lower = medv_mean - 2*mean_se
upper = medv_mean + 2*mean_se
cat("Lower Estimate: ", lower, "Upper Estimate: ", upper)
```

Now we can perform a t-test to find an alternate estimate
```{r}
t.test(Boston[["medv"]])
```

Both estimates are very close to eachoter. We can give a simple point estimate of the median
```{r}
medv_median = median(Boston[["medv"]])
medv_median
```

We use a bootstrap procedure to estimate the standard error with a point estimate given by the mean of the bootstrap distribution
```{r}
medv_median_boot = numeric(N)
for (i in 1:N){
  medv_boot = sample(Boston[["medv"]], n, replace = TRUE)
  medv_median_boot[i] = sd(medv_boot)/sqrt(n)
}
mean(medv_median_boot)
```

The standard error of the median is very similar to that of the mean. We can find the tenth percentile of medv with the quantile function
```{r}
medv_tenth = quantile(Boston[["medv"]], 0.1)
medv_tenth
```

Using the bootstrap method to find the standard error gives
```{r}
medv_tenth_boot = numeric(N)
for (i in 1:N){
  medv_boot = sample(Boston[["medv"]], n, replace = TRUE)
  medv_tenth_boot[i] = quantile(medv_boot, 0.1)
}
sd(medv_tenth_boot)/sqrt(N)
```