---
title: "Exercise 5.8"
output: html_document
---
**(a)** Generating some simulated data
```{r}
set.seed(1)
x=rnorm(100)
y=x-2*x^2+rnorm(100)
```
In this data set the number of samples $n$ is 100 and the standard deviation $\sigma$ is 1. The model is given by $$ P(x) = \frac{1}{\sqrt{2\pi}\sigma}e^{-(x - μ)^2/2 \sigma^2} $$.
**(b)** Creating a scatter plot
```{r}
plot(x,y)
```
The data seems to follow a quadratic trend.

**(c)** Setting a random seed and creating a data frame from the simulated data
```{r}
data=data.frame(x,y)
```

Setting a random seed and performing LOOCV on 4 models of increasing order
```{r}
set.seed(123)
library(boot)
cv.err=rep(0,4)
for (i in 1:4){
  model_i = glm(y~poly(x,i), data=data)
  cv.err[i]=cv.glm(data, model_i)$delta[1]
}
cv.err
```

**(d)** Repeating the previous with a new seed
```{r}
set.seed(321)
library(boot)
cv.err=rep(0,4)
for (i in 1:4){
  model_i = glm(y~poly(x,i), data=data)
  cv.err[i]=cv.glm(data, model_i)$delta[1]
}
cv.err
```

The results for both are identical because the source of randomness has no bearing on a LOOCV due to the fact that regradless of the random seed the LOOCV behavior is deterministic. 

**(e)** The model with the smallest LOOCV error is the quadratic model. I expected this because we know that the underlying data is quadratic itself.

**(f)** Fitting the fourth order model and taking its summary will allow us to see the statistical significance of the coeffecients
```{r}
model = glm(y~poly(x,4),data=data)
summary(model)
```

We can see from the summary that only terms up to second order are significant. This is consistent with the cross validation results as a model containing only the statistically significant terms was the best model.