---
title: "7-1"
output: html_document
---

We begin by loading the data and attaching it for use
```{r}
data = read.csv("Galton.csv", header=TRUE)
attach(data)
```

**(a)** Modifying the heights of females and mothers by multiplying by 1.08
```{r}
data$Mother = 1.08*data$Mother
data[which(Gender=='female'),3] = 1.08*data[which(Gender=='female'),3]
```

Averaging the mother and father heights
```{r}
data["mean"] = (data["Father"]+data["Mother"])/2
```

Fitting a linear model on the mean parent height and child height
```{r}
model = lm(Height~mean, data=data)
summary(model)
```

The slope is less than 1 which indicates that children are less tall than their parents mean height on average. this is the "regression to mediocrity" that Galton noticed.

**(b)** Plotting the residuals
```{r}
plot(model$residuals~model$fitted.values, xlab="Predicted", ylab="Residuals")
abline(h=0)
hist(model$residuals)
```

The scatter plot of residuals appears linear and thus the linearity condition is met. Independence is assumed in the model as heights of children should not be correlated between children only with parents. Plotting  histogram of the residuals reveals that the residuals are approximately normally distributed as is assumed in a linear model with gaussian random noise. Thus, normality is satisfied.

**(d)** Setting a random seed and splitting the data into training and validation sets
```{r}
set.seed(44)
sample = sample.int(n=nrow(data), size=floor(.75*nrow(data)), replace=F)
train = data[sample,]
test = data[-sample,]
```

Estimating test error for models up to degree 3
```{r}
library(boot)
test_error = rep(0,3)
for (i in 1:3){
  model = glm(Height~poly(mean,i), data=train)
  test_error[i] = mean((test$Height - predict.glm(model, test))^2)
}
test_error
```

**(e)** Repeating the analysis above with LOOCV
```{r}
test_error = rep(0,3)
for (i in 1:3){
  model = glm(Height~poly(mean,i), data=data)
  test_error[i] = cv.glm(data, model)$delta[1]
}
test_error
```

**(f)** Finally with $k$-fold cross-validation with $k=5$
```{r}
test_error = rep(0,3)
for (i in 1:3){
  model = glm(Height~poly(mean,i), data=data)
  test_error[i] = cv.glm(data, model, K=5)$delta[1]
}
test_error
```